---
title: "Camera Trap - Représentativité des sites Mt Blanc - Data prep - Bauges"
author: "Jeremy Froidevaux"
date: "2023-11-07"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/jeremyfroidevaux/Library/CloudStorage/GoogleDrive-jfroidevaux@creamontblanc.org/Drive partagés/prototool/herbiland/analyses/cameratrap/cameratrap_landscape")
rm(list=ls()) 
```


## Load libraries
```{r}
library <- c("terra","ggplot2","grid","gridExtra","cowplot","landscapemetrics","data.table","maps",
             "ggsn","dplyr","rworldmap","reshape2","sf","terra","tidyverse","purrr") 
lapply(library, require, character.only = TRUE) 
```

## Data manipulation with raster and shape files
### Open CameraTrap csv file (and landus csv file if needed)
```{r}
camtrap<-fread("data/camerainfoBauges_20231127.csv")

```

### Transformation of latlong in vector and georeferenced data vizualization
```{r}
#All data

pts <- st_as_sf(camtrap,coords=c("long","lat"))
str(pts)
st_crs(pts) <- 2154
st_crs(pts)
st_write(pts, "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Bauges_camtrap/camtrap_locBauges", driver = "ESRI Shapefile")

```

### Plot rasters and camtrap loc
```{r}
dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Bauges_camtrap"  
name.points = "camtrap_locBauges"
loc<-st_read(dsn = dir.posPoints, layer = name.points)
dir.posEnvironmentalLayers = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/PRED/LC_pred_BAUGES_08062023.tif"
orionrast<-rast(dir.posEnvironmentalLayers) 
orionrast
loc
plot(orionrast)
points(loc)

```


## Data extraction
### Part 1: Extract landuse proportion from ORION at local and landscape scales

```{r}
source("function/ExtractBufferRaster_KB_JF.r") 

extractLandUseRaster( dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Bauges_camtrap",  
                                name.points = "camtrap_locBauges",
                                bw = c(20,250),
                                id = "Station",  
                                dir.posEnvironmentalLayers = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/PRED/LC_pred_BAUGES_08062023.tif",  
                                WorkingDirectory = "/Users/jeremyfroidevaux/Documents/Herbiland/CameraTrap/Analysis/Landcover_output/") 

```

### Part 2: Extract data on landscape configurational and compositional diversity using ORION

```{r}
dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Bauges_camtrap"  
name.points = "camtrap_locBauges"
loc<-st_read(dsn = dir.posPoints, layer = name.points)
dir.posEnvironmentalLayers = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/PRED/LC_pred_BAUGES_08062023.tif"
orionrast<-rast(dir.posEnvironmentalLayers) 
orionrast
loc
plot(orionrast)
points(loc)
list<-list_lsm() #Choose landscape metric 

plm20 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$Station,  
                    size = 20, level= "landscape",metric=c("shdi","ed")) 
plm250 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$Station,  
                    size = 250, level= "landscape",metric=c("shdi","ed")) 
pcm20 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$Station,  
                    size = 20, level= "class",metric=c("ed")) 
pcm250 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$Station,  
                    size = 250, level= "class",metric=c("ed")) 

plm20$scale<-20 
plm250$scale<-250
pcm20$scale<-20 
pcm250$scale<-250

pcm20sel<-pcm20[pcm20$class %in% c("5"), ]
pcm250sel<-pcm250[pcm250$class %in% c("5"), ]
pcm20sel$metric[pcm20sel$metric == 'ed'] <- 'edl'
pcm250sel$metric[pcm250sel$metric == 'ed'] <- 'edl'

df_list <- list(plm20,plm250,pcm20sel,pcm250sel)
mplm<-Reduce(function(x, y) merge(x, y, all=TRUE), df_list)

div <- dcast(mplm, plot_id + scale~ metric, fun.aggregate = sum)

#ggplot(subset(rmplm, scale %in% 50))+
#  geom_histogram(aes(x=ed))

div20<-subset(div,scale=="20")
div20<-div20 %>% rename_with(~paste0(.x,"20"), c(ed,edl,shdi ))

div250<-subset(div,scale=="250")
div250<-div250 %>% rename_with(~paste0(.x,"250"), c(ed,edl,shdi ))

df_list <- list(div20, div250)
divf<-df_list %>% reduce(full_join, by='plot_id')
divf<-divf %>% dplyr::select(-contains("scale"))
colnames(divf)[colnames(divf) == "plot_id"] <- "Station"

write.csv(divf, "output/landscapeMetric_outputBauges.csv", row.names=FALSE) 

```

### Part 4: DEM
#### Crop DEM to study area
```{r}
dir.dem = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Topo/DEM_Alps/DEM_ALPS1.tif"
dem<-rast(dir.dem)
demproj<-project(dem, "EPSG:2154") #EPSG:2154 pour Lambert93
demproj
plot(demproj,axes = T,legend=FALSE)

dir.posEnvironmentalLayers = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/PRED/LC_pred_BAUGES_08062023.tif"
orionm<-rast(dir.posEnvironmentalLayers)
demBauges<-crop(demproj, orionm, snap="near", extend=FALSE,mask=TRUE)

dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Bauges_camtrap"  
name.points = "camtrap_locBauges"
loc<-st_read(dsn = dir.posPoints, layer = name.points)
plot(demBauges,axes = T,legend=T)
points(loc)

writeRaster(demBauges,file="/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Topo/DEM_Alps/DEM_Bauges.tif")
```

#### Extract elevation, slope and aspect 
```{r}
dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Bauges_camtrap"  
name.points = "camtrap_locBauges"
loc<-st_read(dsn = dir.posPoints, layer = name.points)
#loc <- st_transform(loc,crs=32632)
loc
dir.dem = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Topo/DEM_Alps/DEM_Bauges.tif"
dem<-rast(dir.dem)
slope <- terrain(dem, "slope",neighbors = 8, unit = "degrees")
aspect <- terrain(dem, 'aspect', neighbors = 8, unit = "degrees")

plot(slope)
points(loc)

elevationCT<-extract(dem, loc)
slopeCT<-extract(slope, loc)
aspectCT<-extract(aspect, loc)

demCT<-bind_cols(elevationCT,slopeCT,aspectCT)
demCT<-data.frame(Station=camtrap$Station, demCT)
demCT<-demCT[ -c(2,4,6) ]
colnames(demCT)[2] ="elevation"
write.csv(demCT,"output/dem_outputBauges.csv")
```
