---
title: "Camera Trap - Représentativité des sites Mt Blanc - Data prep"
author: "Jeremy Froidevaux"
date: "2023-11-07"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/jeremyfroidevaux/Library/CloudStorage/GoogleDrive-jfroidevaux@creamontblanc.org/Drive partagés/prototool/herbiland/analyses/cameratrap/cameratrap_landscape")
rm(list=ls()) 
```


## Load libraries
```{r}
library <- c("terra","ggplot2","grid","gridExtra","cowplot","landscapemetrics","data.table","maps",
             "ggsn","dplyr","rworldmap","reshape2","sf","terra","tidyverse","purrr") 
lapply(library, require, character.only = TRUE) 
```

## Data manipulation with raster and shape files
### Open CameraTrap csv file (and landus csv file if needed)
```{r}
camtrap<-fread("data/camerainfo_20231116.csv")
camtrap<-camtrap[camtrap$running=="Y",] #camtrap active
camtrapVCM<-camtrap[camtrap$lat>45.85,] #ComCom Chamonix
camtrapVCM$milieu<-gsub("lande sur crete", "lande",camtrapVCM$milieu)
camtrapVCM<-camtrapVCM[!grepl("campagnol", camtrapVCM$Station),] #remove site with campagnol
camtrapVCM_orchamp<-camtrapVCM[grepl("Y", camtrapVCM$ORCHAMP),] #focuis on orchamlp sites

#write.csv(camtrapVCM,"data/camerainfo_20231113_sel.csv")
#write.csv(camtrapVCM_orchamp,"data/camerainfo_orchamp_20231113_sel.csv")

#camtrapPMB<-camtrap[camtrap$lat<45.85,] #ComCom vallée du mont blanc

#landus<-fread("/Users/jeremyfroidevaux/Documents/Herbiland/CameraTrap/landus_20231109.csv")
#summary(factor(landus$ref_project))
```

### Transformation of latlong in vector and georeferenced data vizualization
```{r}
#All data
lonlat<-cbind(camtrapVCM$long,camtrapVCM$lat)
crdref <- "+proj=longlat +datum=WGS84"
pts <- vect(lonlat, crs=crdref)
pts
pts$id<-camtrapVCM$Station
pts$orchamp<-camtrapVCM$ORCHAMP
class(pts)
plot(pts)
pts
geom(pts)
outfile <- "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/CamTrap_location/camtrap_loc.shp"
writeVector(pts, outfile, overwrite=TRUE)

#Focus on orchamp
lonlat<-cbind(camtrapVCM_orchamp$long,camtrapVCM_orchamp$lat)
crdref <- "+proj=longlat +datum=WGS84"
pts <- vect(lonlat, crs=crdref)
pts
pts$id<-camtrapVCM_orchamp$Station
class(pts)
plot(pts)
pts
geom(pts)
outfile <- "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/CamTrap_location/camtrap_orchamp_loc.shp"
writeVector(pts, outfile, overwrite=TRUE)

```

### Plot & combine rasters
```{r}

dir.posEnvironmentalLayers1 = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/LC_pred_CCVCMB_2806023_FINAL.tif"
orionrast1<-rast(dir.posEnvironmentalLayers1)
dir.posEnvironmentalLayers2 = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/LC_pred_CCPMB_2806023_FINAL.tif"
orionrast2<-rast(dir.posEnvironmentalLayers2) 
#summary(factor(values(orionrast1)))
#summary(factor(values(orionrast2)))

plot(orionrast1)
plot(orionrast2)
s <- sprc(orionrast1, orionrast2)

orionrastm <- terra::merge(s)
summary(factor(values(orionrastm)))

writeRaster(orionrastm,file="/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/merge_CCPMB_CCVCMB_280623_FINAL.tif")
dir.orionmerge = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/merge_CCPMB_CCVCMB_280623_FINAL.tif"
orionrastm<-rast(dir.orionmerge)
orionrastm
values(orionrastm)
summary(factor(values(orionrastm)))
```


## Data extraction
### Part 1: Extract landuse proportion from ORION at local and landscape scales

```{r}
root <- "/Users/jeremyfroidevaux/Documents/Herbiland/CameraTrap/Analysis/Functions/" 
source(paste(root,"ExtractBufferRaster_KB_JF.r",sep="")) 

extractLandUseRaster( dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/CamTrap_location",  
                                name.points = "camtrap_loc",
                                bw = c(20,250), #ajouter 20 m - 250 m  
                                id = "id",  
                                dir.posEnvironmentalLayers = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/LC_pred_CCVCMB_2806023_FINAL.tif",  
                                WorkingDirectory = "/Users/jeremyfroidevaux/Documents/Herbiland/CameraTrap/Analysis/Landcover_output/") 

```

### Part 2: Extract data on landscape configurational and compositional diversity using ORION

```{r}
dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/CamTrap_location"  
name.points = "camtrap_loc"
loc<-st_read(dsn = dir.posPoints, layer = name.points)
loc <- st_transform(loc,crs=32632)
dir.posEnvironmentalLayers = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/LC_pred_CCVCMB_2806023_FINAL.tif"
orionrast<-rast(dir.posEnvironmentalLayers) 
orionrast
loc
plot(orionrast)
points(loc)
list<-list_lsm() #Choose landscape metric 

plm20 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$id,  
                    size = 20, level= "landscape",metric=c("shdi","ed")) 
plm250 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$id,  
                    size = 250, level= "landscape",metric=c("shdi","ed")) 
pcm20 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$id,  
                    size = 20, level= "class",metric=c("ed")) 
pcm250 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$id,  
                    size = 250, level= "class",metric=c("ed")) 

plm20$scale<-20 
plm250$scale<-250
pcm20$scale<-20 
pcm250$scale<-250

pcm20sel<-pcm20[pcm20$class %in% c("5"), ]
pcm250sel<-pcm250[pcm250$class %in% c("5"), ]
pcm20sel$metric[pcm20sel$metric == 'ed'] <- 'edl'
pcm250sel$metric[pcm250sel$metric == 'ed'] <- 'edl'

df_list <- list(plm20,plm250,pcm20sel,pcm250sel)
mplm<-Reduce(function(x, y) merge(x, y, all=TRUE), df_list)

div <- dcast(mplm, plot_id + scale~ metric, fun.aggregate = sum)

#ggplot(subset(rmplm, scale %in% 50))+
#  geom_histogram(aes(x=ed))

div20<-subset(div,scale=="20")
div20<-div20 %>% rename_with(~paste0(.x,"20"), c(ed,edl,shdi ))

div250<-subset(div,scale=="250")
div250<-div250 %>% rename_with(~paste0(.x,"250"), c(ed,edl,shdi ))

df_list <- list(div20, div250)
divf<-df_list %>% reduce(full_join, by='plot_id')
divf<-divf %>% dplyr::select(-contains("scale"))
colnames(divf)[colnames(divf) == "plot_id"] <- "id"

write.csv(divf, "output/landscapeMetric_outputVCMsel.csv", row.names=FALSE) 

```

### Part 4: DEM
#### Crop DEM to study area
```{r}
dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/CamTrap_location"  
name.points = "camtrap_loc"
dir.dem = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Topo/DEM_Alps/DEM_ALPS1.tif"
loc<-st_read(dsn = dir.posPoints, layer = name.points)
loc <- st_transform(loc,crs=32632)
loc
dem<-rast(dir.dem)
demproj<-project(dem, "EPSG:32632")
demproj
plot(demproj,axes = T,legend=FALSE)
points(loc)

dir.posEnvironmentalLayers = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/merge_CCPMB_CCVCMB_280623_FINAL.tif"
orionm<-rast(dir.posEnvironmentalLayers)
demMB<-crop(demproj, orionm, snap="near", extend=FALSE,mask=TRUE)

writeRaster(demMB,file="/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Topo/DEM_Alps/DEM_MB.tif")
```

#### Extract elevation, slope and aspect 
```{r}
dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/CamTrap_location"  
name.points = "camtrap_loc"
loc<-st_read(dsn = dir.posPoints, layer = name.points)
loc <- st_transform(loc,crs=32632)
loc
dir.dem = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Topo/DEM_Alps/DEM_MB.tif"
dem<-rast(dir.dem)
slope <- terrain(dem, "slope",neighbors = 8, unit = "degrees")
aspect <- terrain(dem, 'aspect', neighbors = 8, unit = "degrees")

elevationCT<-extract(dem, loc)
slopeCT<-extract(slope, loc)
aspectCT<-extract(aspect, loc)

demCT<-bind_cols(elevationCT,slopeCT,aspectCT)
demCT<-data.frame(id=camtrapVCM$Station, demCT)
demCT<-demCT[ -c(2,4,6) ]
colnames(demCT)[2] ="elevation"
plot(camtrapVCM$alt,demCT$elevation)
write.csv(demCT,"o
          utput/dem_outputCVMsel.csv")
```
